### 2019-07-07 juliacon2019proceedings/paper/header.tex
Paper upload

% **************GENERATED FILE, DO NOT EDIT**************

\title{GlobalSearchRegression.jl: \\ Building bridges between Machine Learning and Econometrics in Fat-Data scenarios}

\author[1, 3, 4]{Demian Panigo}
\author[1, 2]{Plabo Glüzmann}
\author[1, 5]{Esteban Mocskos}
\author[3]{Adan Mauri Ungaro}
\author[3]{Valentín Mari}
\author[3, 4]{Nicolás Monzón}

\affil[1]{National Scientific and Technical Research Council (CONICET)}
\affil[2]{Center For Distributive, Labor and Social Studies (CEDLAS)}
\affil[3]{National University of La Plata}
\affil[4]{National University of Avellaneda}
\affil[5]{National University of Buenos Aires}

\keywords{Julia, Parallel computing, Econometrics, all-subset regression, Machine Learning, Fat-Data}